---
title: "Text Mining and Sentiment Analysis"
author: "John Cruz"
date: "2023-04-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In Text Mining with R, [Chapter 2](https://www.tidytextmining.com/sentiment.html) looks at sentiment analysis. The authors provides an example using the text of Jane Austen’s six completed, published novels from the *janeaustenr* library. All the code is originally credited to the authors, unless otherwise noted.

---

## Required Libraries

```{r library, message=FALSE}
library(tidyverse)
library(tidytext)
library(janeaustenr)
library(stringr)
library(jsonlite)
library(glue)
library(lubridate)
```

---

## Tidy Up Jane Austen's Work

The authors take the text of the novels and converts the text to the tidy format using *unnest_tokens()*. They also create other columns to keep track of which line and chapter of the book each word comes from. 

```{r example-1}
tidy_books <- 
  austen_books() |> 
  group_by(book) |> 
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", ignore_case = TRUE)))) |> 
  ungroup() |> 
  unnest_tokens(word, text)

knitr::kable(head(tidy_books), caption = "Brief View of Tokenized Words")
```

## Determining Overall Sentiment

Next, count up how many positive and negative words there are in defined sections of each book, along with a net sentiment score. They define an index here to keep track of where they are in the narrative. The index counts up sections of 80 lines of text.

```{r example-2, message=FALSE, warning=FALSE}
jane_austen_sentiment <- 
  tidy_books |>
  inner_join(get_sentiments("bing")) |>
  count(book, index = linenumber %/% 80, sentiment) |>
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) |> 
  mutate(sentiment = positive - negative)

knitr::kable(head(jane_austen_sentiment), caption = "Brief View of Sentiment Scores by Indexing")
```

## Visualizing Sentiment throughout each Novel

Finally, plot how each novel changes toward more positive or negative sentiment over the trajectory of the story.

```{r example-3}
jane_austen_sentiment |> 
  ggplot(aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

## Which Chapter has the Most Negative Words?

The authors also provide a proportion table to determine which chapter has the most negative words in each book. 

```{r example-4, warning=FALSE, message=FALSE}
bingnegative <- 
  get_sentiments("bing") |> 
  filter(sentiment == "negative")

wordcounts <- tidy_books |>
  group_by(book, chapter) |>
  summarize(words = n())

ratio_tbl <-
  tidy_books |>
  semi_join(bingnegative) |>
  group_by(book, chapter) |>
  summarize(negativewords = n()) |>
  left_join(wordcounts, by = c("book", "chapter")) |>
  mutate(ratio = negativewords/words) |>
  filter(chapter != 0) |>
  slice_max(ratio, n = 1) |> 
  ungroup()

knitr::kable(ratio_tbl)
```

---

**Note:** All work from this point forward has been performed by me. 

## Corpus: NY Times Articles

March 2023

## Connect to NY Times API

```{r example}
nyt_key <- rstudioapi::askForPassword('Enter NY Times API Key')

api_cnxn <- 
  fromJSON(glue("https://api.nytimes.com/svc/archive/v1/2023/3.json?api-key={nyt_key}"), flatten = TRUE)

ny_times <- 
    as.data.frame(api_cnxn) |> 
    janitor::clean_names()

write_csv(ny_times, 'ny_times.csv')
```

## Tidy Up Article Data

```{r example}
lead_paragraph <-
  ny_times |> 
  select(response_docs_pub_date, response_docs_lead_paragraph, response_docs_section_name) |> 
  mutate(response_docs_pub_date = str_extract(response_docs_pub_date, "[:graph:]*(?=\\+)")) |> 
  rename(pub_date = response_docs_pub_date, lead_paragraph = response_docs_lead_paragraph, section = response_docs_section_name)

lead_paragraph$pub_date <- 
  lead_paragraph$pub_date |> 
  ymd_hms()
```

## Title

```{r example}
```

## Title

```{r example}
```

## Title

```{r example}
```

## Title

```{r example}
```






