---
title: "Sentiment Analysis"
author: "John Cruz"
date: "2023-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In Text Mining with R, Chapter 2 looks at Sentiment Analysis.  In this assignment, you should start by getting the primary example code from chapter 2 working in an R Markdown document.  You should provide a citation to this base code.  You’re then asked to extend the code in two ways:
Work with a different corpus of your choosing, and
Incorporate at least one additional sentiment lexicon (possibly from another R package that you’ve found through research).

https://www.tidytextmining.com/sentiment.html

## Required Libraries

```{r library}
library(tidyverse)
library(tidytext)
library(janeaustenr)
library(stringr)
```

## Including Plots

You can also embed plots, for example:

```{r example}
text <- c("Because I could not stop for Death -",
          "He kindly stopped for me -",
          "The Carriage held but just Ourselves -",
          "and Immortality")

text_df <- tibble(line = 1:4, text = text)

text_df |> 
  unnest_tokens(word, text)
```

```{r example-2}
original_books <- austen_books() |> 
  group_by(book) |> 
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", ignore_case = TRUE)))) |> 
  ungroup()

tidy_books <- original_books |> 
  unnest_tokens(word, text)

# remove stopwords
tidy_books <- tidy_books %>%
  anti_join(stop_words)

tidy_books %>%
  count(word, sort = TRUE) 
```











